---
editor: 
  markdown: 
    wrap: sentence
---

## Introduction

## Part 1: Data Design and Implementation

In this report, we assume that the E-commerce platform mainly sells IT products in the United Kingdom and does not have an international shipment policy.
Furthermore, company policy is that product only has one promotion during a year, “Product_ID” is identified based on “Supplier_ID”, which means supplier only has one “Product_ID” for one product.
Also, money collected by e-commerce company is distributed later to supplier.

There are 7 entities:

1.Buyer: Represents the customers who create an account to place orders on the online shopping platform.

2.Supplier: Represents the sellers who create an account to sell products on the platform

3.Product: Represents the set of products available for purchase on the platform.

4.Transaction: Represents product orders placed by customers.

5.Promotion: Represents product promotions to attract willing customers.

6.Shipper: Represents Shipper information.

7.Reviews: Represents the reviews made by the customer once the product is received.

\[picture\]

## Part 2: Data generation

## Part 3:


```{r}
unique(address_data$county)
```


## Part 4:

```{r}
db_conn <- dbConnect(RSQLite::SQLite(), "database/ecommerce.db")
```


1) Products With Highest Sales

```{r}
top_5_products_sales <- dbGetQuery(db_conn, "SELECT
    od.product_id,
    p.product_name,
    SUM(p.price) AS total_sales
FROM
    order_details od
JOIN
    product p ON od.product_id = p.product_id
GROUP BY
    od.product_id,
    p.product_name
ORDER BY
    total_sales DESC
LIMIT
    5;
")

# Viewing the results
print(top_5_products_sales)

```


