---
title: "Q4Test"
output: html_document
date: "2024-03-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages(c("DBI", "RSQLite", "dplyr","ggplot2"))

library(DBI)
library(RSQLite)
library(dplyr)
library(ggplot2)
```

```{r}
db_conn <- dbConnect(RSQLite::SQLite(), "database/ecommerce.db")
```

1) Products With Highest Sales
```{r}
top_5_products_sales <- dbGetQuery(db_conn, "SELECT
    od.product_id,
    p.product_name,
    SUM(p.price) AS total_sales
FROM
    order_details od
JOIN
    product p ON od.product_id = p.product_id
GROUP BY
    od.product_id,
    p.product_name
ORDER BY
    total_sales DESC
LIMIT
    5;
")
```


```{r}
ggplot(top_5_products_sales, aes(x = reorder(product_name, total_sales), y = total_sales, fill = product_name)) +
  geom_col() + 
  labs(x = "Product Name", y = "Total Sales") +
  theme_minimal() +
  scale_fill_viridis_d() +
  coord_flip() + 
  ggtitle("Top 5 Products by Sales") +
  theme(legend.position = "none") 
```


2) Promotion for top products
```{r}
promotion_data <- dbGetQuery(db_conn, "SELECT * FROM promotion;")
top_5_products_promotion <- dbGetQuery(db_conn, "SELECT
    od.product_id,
    p.product_name,
    p.promotion_id,
    pm.status,
    SUM(p.price) AS total_sales
FROM
    order_details od
JOIN
    product p ON od.product_id = p.product_id
JOIN 
    promotion pm ON p.promotion_id = pm.promotion_id
GROUP BY
    od.product_id,
    p.product_name
ORDER BY
    total_sales DESC
LIMIT
    5;
")

print(top_5_products_promotion)

```

```{r}
ggplot(top_5_products_promotion, aes(x = reorder(product_name, total_sales), y = total_sales, fill = status)) +
  geom_col() + 
  labs(title = "Top 5 Products by Sales", x = "Product Name", y = "Total Sales") +
  theme_minimal() +
  scale_fill_viridis_d() +
  coord_flip() + 
  scale_fill_discrete(name = "Promotion Status") + 
  theme(legend.position = "right") 
```

3) Number of days promotions are running (Not Working)
```{r}
no_of_days_promotions <- mutate(promotion_data,
                         start_date = as.Date(start_date,format= "%d/%m/%Y"),
                         end_date = as.Date(end_date,format= "%d/%m/%Y"),
                         num_days_promotion = end_date - start_date)
no_of_days_promotions <- no_of_days_promotions %>%
  arrange(desc(num_days_promotion))

no_of_days_promotions <- no_of_days_promotions %>% 
  group_by(num_days_promotion) %>% 
  summarise(num_promotions = n()) %>%
  arrange((num_days_promotion))
print(no_of_days_promotions)

```
```{r}
ggplot(no_of_days_promotions, aes(x = num_days_promotion, y = num_promotions)) +
  geom_col() + 
  labs(title = "Number of Promotions running across each different duration", x = "Number of days promotion is running", y = "Number of promotions") +
  theme_minimal() +
  scale_fill_viridis_d() +
  theme(legend.position = "none")
```

4) Sales Trends
```{r}
order_details <- dbGetQuery(db_conn,"SELECT * FROM order_details")
order_details <- mutate(order_details,
                        order_dates = as.Date(order_dates,format= "%d/%m/%Y"))
product <- dbGetQuery(db_conn,"SELECT * FROM product")


# Viewing the results
sales_trend_data <- inner_join(order_details, product, 
           by = c("product_id")) 

sales_trend_data <- sales_trend_data %>%
  group_by(order_dates) %>% 
  summarise(total_sales = sum(price))

```
```{r}
ggplot(sales_trend_data, aes(x = order_dates, y = total_sales)) +
  geom_line() + 
  labs(title = "Sales Trend", x = "Date", y = "Sales") +
  theme_minimal() +
  scale_fill_viridis_d() +
  theme(legend.position = "none")
```
5) Products With Highest Review
```{r}
top_5_products_reviews <- dbGetQuery(db_conn, "SELECT
    r.product_id,
    p.product_name,
    AVG(r.review_score) AS avg_review_score
FROM
    review r
JOIN
    product p ON r.product_id = p.product_id
GROUP BY
    r.product_id,
    p.product_name
ORDER BY
    avg_review_score DESC
LIMIT
    5;
")

# Viewing the results
print(top_5_products_reviews)
```

```{r}
ggplot(top_5_products_reviews, aes(x = reorder(product_name, avg_review_score), y = avg_review_score, fill = product_name)) +
  geom_col(show.legend = FALSE) + 
  coord_flip() + 
  labs(title = "Top 5 Products by Average Review Score", x = "Product Name", y = "Average Review Score") +
  theme_minimal() +
  scale_fill_viridis_d() + 
  theme(plot.title = element_text(hjust = 0.5)) 

```


6) Products With Lowest Review

```{r}
lowest_5_products_reviews <- dbGetQuery(db_conn, "SELECT
    p.product_id,
    p.product_name,
    AVG(r.review_score) AS average_review_score
FROM
    review r
JOIN
    product p ON r.product_id = p.product_id
GROUP BY
    p.product_id,
    p.product_name
ORDER BY
    average_review_score ASC
LIMIT
    5;
")

# Viewing the results
print(lowest_5_products_reviews)
```

```{r}
ggplot(lowest_5_products_reviews, aes(x = reorder(product_name, -average_review_score), y = average_review_score)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Products with Lowest Review Scores", x = "Product", y = "Average Review Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


7) Cities With Lowest Review

```{r}

lowest_5_city_reviews <- dbGetQuery(db_conn, "SELECT
    a.city,
    AVG(r.review_score) AS average_review_score
FROM
    review r
JOIN
    buyer b ON r.buyer_id = b.buyer_id
JOIN
    address a ON b.address_id = a.address_id
GROUP BY
    a.city
ORDER BY
    average_review_score ASC
LIMIT
    5;
")

# Viewing the results
print(lowest_5_city_reviews)

```


```{r}
ggplot(lowest_5_city_reviews, aes(x = reorder(city, average_review_score), y = average_review_score)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Cities with Lowest Review Scores", x = "City", y = "Average Review Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

8) Cities With Highest Review

```{r}

highest_5_city_reviews <- dbGetQuery(db_conn, "SELECT
    a.city,
    AVG(r.review_score) AS average_review_score
FROM
    review r
JOIN
    buyer b ON r.buyer_id = b.buyer_id
JOIN
    address a ON b.address_id = a.address_id
GROUP BY
    a.city
ORDER BY
    average_review_score DESC
LIMIT
    5;
")

# Viewing the results
print(highest_5_city_reviews)

```

```{r}
ggplot(highest_5_city_reviews, aes(x = reorder(city, -average_review_score), y = average_review_score)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Cities with Highest Review Scores", x = "City", y = "Average Review Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

9) Estimate the CLV based on historical transactions to identify your most valuable customers.

```{r}
customer_lifetime_value <- dbGetQuery(db_conn, "SELECT
    buyer_id,
    SUM(p.price) AS total_spent,
    COUNT(distinct order_id) AS number_of_orders,
    (SUM(p.price) / COUNT(distinct order_id)) AS average_order_value
FROM
    order_details od
JOIN
    product p ON od.product_id = p.product_id
GROUP BY
    buyer_id
ORDER BY
    total_spent DESC;
")

# Viewing the results
print(customer_lifetime_value)

```

```{r}
top_n_buyers <- head(customer_lifetime_value, 10)  
# Change 10 to the desired number of top buyers to display

ggplot(top_n_buyers, aes(x = reorder(buyer_id, -total_spent), y = total_spent)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Top Buyers by Customer Lifetime Value", x = "Buyer ID", y = "Total Spent") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


10) Effective of Promotions

```{r}
effective_promotions <- dbGetQuery(db_conn, "SELECT
    promo.promotion_type,
    COUNT(*) AS number_of_sales,
    SUM(prod.price) AS total_revenue
FROM
    order_details od
JOIN
    product prod ON od.product_id = prod.product_id
JOIN
    promotion promo ON prod.promotion_id = promo.promotion_id
WHERE
    promo.status = 'active'
GROUP BY
    promo.promotion_type
ORDER BY
    total_revenue DESC;;
")

# Viewing the results
print(effective_promotions)

```

```{r}
ggplot(effective_promotions, aes(x = reorder(promotion_type, -total_revenue), y = total_revenue)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Effectiveness of Promotions", x = "Promotion Type", y = "Total Revenue") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

11) Most Reviewed Products

```{r}
most_reviewed_products <- dbGetQuery(db_conn,"SELECT
    p.product_name,
    COUNT(*) AS review_count,
    AVG(r.review_score) AS avg_review_score
FROM
    review r
JOIN
    product p ON r.product_id = p.product_id
GROUP BY
    p.product_name
ORDER BY
    review_count DESC
LIMIT 10;
")

# Viewing the results
print(most_reviewed_products)

```


```{r}
ggplot(most_reviewed_products, aes(x = reorder(product_name, -review_count), y = review_count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Most Reviewed Products", x = "Product Name", y = "Number of Reviews") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
The high number of reviews for Touch Desk Lamp and Ultra-Light Smart Speaker indicates strong customer engagement with these products. More customers are actively providing feedback for these products, which can be valuable for understanding product satisfaction and areas for improvement.


12) Cities With Highest Sales.
```{r}

cities_highest_sales <- dbGetQuery(db_conn,"SELECT
    a.city,
    SUM(p.price) AS total_sales
FROM
    order_details od
JOIN
    product p ON od.product_id = p.product_id
JOIN
    buyer b ON od.buyer_id = b.buyer_id
JOIN
    address a ON b.address_id = a.address_id
GROUP BY
    a.city
ORDER BY
    total_sales DESC
LIMIT
    5;
")

# Viewing the results
print(cities_highest_sales)

```


```{r}
ggplot(cities_highest_sales, aes(x = city, y = total_sales)) +
  geom_bar(stat = "identity", fill = "skyblue", width = 0.5) +
  labs(title = "City with Highest Sales", x = "City", y = "Total Sales") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```









